Investigating magnetic dynamo evolution with TESS field dwarfs

### 2017-10-05 ms.tex
mag limit

Some outstanding mysteries regarding the nature of magnetic braking have been
revealed, but thus-far unanswered by, \kepler.
Firstly, a mysterious gap in the rotation period distribution of \Kepler\
dwarfs requires either a sharp transition in magnetic dynamo geometry or a gap
in the local star formation history as an explanation \citep{mcquillan2014,
davenport2017}.
Secondly, a transitioning magnetic dynamo appears to be responsible for
inefficient magnetic braking at old ages in Solar-mass stars.
{\bf We propose to help answer these questions by providing a catalog of \TESS\
rotation periods for FFI and CTL stars.}
% We also intend to use gyrochronology in the cases where it {\it is} applicable
% to learn about the evolution of exoplanets across the galaxy.

@@

The full set of available \TESS\ FFI and two-minute cadence light curves will
be exceedingly large and we intend to prioritize certain groups of stars for
period analysis, calculating a simple (and computationally fast) Lomb-Scargle
periodogram for the low priority objects and running a full probabilistic
period analysis on the high priority targets.
High priority targets include: comoving pairs, stars in the CVZs (especially
those with spectra), stars in TGAS, stars at large distances, plus any
potential asteroseismic dwarfs and exoplanet hosts.
% Of these, the stars that also lie in the CVZs as well as the identified \Gaia\
% comoving pairs will be our highest priority.
% Our next priority will be stars in the CVZs, with spectra, that do not appear
% in TGAS, followed by those without spectra.
% M dwarfs rotation periods in the CVZz will allow us to study the rotation
% period bimodality at lower masses, since the bimodality appears at longer
% rotation periods for lower mass stars.
% We will then focus on measuring rotation periods for a sample of stars that is
% representative of the full set of FFI and CTL targets in order to enable
% exoplanet population studies.
% An large amount of computational resources are available to us, with access to
% the Columbia supercomputer as well as facilities at the Flatiron Institute.
% We expect that each target will require around one minute of computation time
% and have access to around 10,000 computing hours.
% We therefore expect to analyze around 60,000 light curves.

@@

\paragraph{Expected Impact}
Initially we will provide rotation period posterior probability density
functions and best fit estimates for around 30,000 high priority two-minute
cadence and FFI targets.
We will then produce a catalog of rotation periods for lower priority stars
using a combination of Lomb-Scargle periodogram and autocorrelation
techniques.

\paragraph{Budget Justification}
PI Angus intends to use the budget to employ a student for 3-4 months.
In this time the student will assist in extracting light curves from the FFIs,
measuring rotation periods and building a rotation period catalog.
The student will also be involved in the scientific project of their choosing:
either the rotation period bimodality, gyrochronology of low-mass stars, or
gyrochronology of comoving pairs.

\paragraph{References}

>>>>

Some outstanding mysteries regarding the nature of magnetic braking have been
revealed, but thus-far unanswered by, \kepler.
Firstly, a mysterious gap in the rotation period distribution of \Kepler\
dwarfs requires either a sharp transition in magnetic dynamo geometry or a gap
in the local star formation history as an explanation \citep{mcquillan2014,
davenport2017}.
Secondly, a transitioning magnetic dynamo appears to be responsible for
inefficient magnetic braking at old ages in Solar-mass stars.
{\bf We propose to help answer these questions by providing a catalog of
rotation periods for \TESS\ Full Frame Image (FFI) stars.}
% We also intend to use gyrochronology in the cases where it {\it is} applicable
% to learn about the evolution of exoplanets across the galaxy.

@@

The full set of available \TESS\ FFI and two-minute cadence light curves will
be exceedingly large and we intend to prioritize certain groups of stars for
period analysis, calculating a simple (and computationally fast) Lomb-Scargle
periodogram for the low priority objects and running a full probabilistic
period analysis on the high priority targets.
High priority targets include: comoving pairs, stars in the CVZs (especially
those with spectra), stars in TGAS, stars at large distances, plus any
potential asteroseismic dwarfs and exoplanet hosts.
Based on the expected photometric precision of \tess\ \citep{sullivan2015}, we
expect to be able to measure rotation periods for FFI stars down to 16th
magnitude.
We note, however that not all stars will have measurable rotation periods,
even if they are bright enough --- not all stars show rotational variability
in their light curves either because they are pole-on or because they are
inactive.
% Of these, the stars that also lie in the CVZs as well as the identified \Gaia\
% comoving pairs will be our highest priority.
% Our next priority will be stars in the CVZs, with spectra, that do not appear
% in TGAS, followed by those without spectra.
% M dwarfs rotation periods in the CVZz will allow us to study the rotation
% period bimodality at lower masses, since the bimodality appears at longer
% rotation periods for lower mass stars.
% We will then focus on measuring rotation periods for a sample of stars that is
% representative of the full set of FFI and CTL targets in order to enable
% exoplanet population studies.
% An large amount of computational resources are available to us, with access to
% the Columbia supercomputer as well as facilities at the Flatiron Institute.
% We expect that each target will require around one minute of computation time
% and have access to around 10,000 computing hours.
% We therefore expect to analyze around 60,000 light curves.

@@

\paragraph{Expected Impact}
Initially we will provide rotation period posterior probability density
functions and best fit estimates for around 30,000 high priority two-minute
cadence and FFI targets.
We will then produce a catalog of rotation periods for lower priority stars
using a combination of Lomb-Scargle periodogram and autocorrelation
techniques.
Finally, we will release a catalog of rotation periods for two-minute cadence
targets.
We will also release an open source python package for measuring rotation
periods of both FFI and two-minute cadence targets.

\paragraph{Budget Justification}
PI Angus intends to use the budget to employ a student for 3-4 months.
In this time the student will assist in extracting light curves from the FFIs,
measuring rotation periods and building a rotation period catalog.
The student will also be involved in the scientific project of their choosing:
either the rotation period bimodality, gyrochronology of low-mass stars, or
gyrochronology of comoving pairs.
The remaining budget will be used for traveling to meet with other Co-Is in
order to perform this analysis.
\paragraph{References}
